Parallel Computing for Constrained Global Optimization Problems \thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}

### 2018-06-14 2018/Journal of Supercomputing/JoS_Paper.tex
Some new text has been included

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{multirow}
%
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
 \journalname{The Journal of Supercomputing}
%
\begin{document}

@@

The aim of this paper is to examine parallel methods for solving global optimization problems with non-convex constraints. The objective function and constraints are assumed to satisfy the Lipschitz condition with a priori unknown Lipschitz constants. The analytical form of the problem's functions may also be unknown, i.e. they can be set by an algorithm computing their values at various points within the search domain (the so-called ``black-box'' functions). It is also supposed that even a single computation of the function value can be time-consuming, since it requires performing numerical simulation in applied problems. 
%These assumptions are typical for many approaches to building parallel algorithms for unconstrained global optimization \cite{Jones2001,Paulavicius2011,Evtushenko2013}.

@@

Among the examples of such problems are the optimal control problems described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2011,Balandin2017}. It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e. every eigenvalue of $A$ has a strictly negative real part. Otherwise, the value of the criteria is indeterminate.

@@

@@

@@

%At every iteration of the method, the process with the index $s, 0\leq s < S,$ performs $p$ trials in parallel at the points $x^{s+iS}, 0\leq i < p$. At that, each process stores all $S\cdot p$ points, and an attribute indicating whether this point is blocked by another process or not is stored for each point. Let us remind that the point is blocked if the process starts the execution of a trial at this point.

%At every iteration of the algorithm, operating within the $s$-th process, determines the coordinates of $p$ ``its own'' trial points. Then, the interchange of the coordinates of images of the trial points $y^{s+iS},\; 0\leq i<p,\; 0\leq s< S$ is performed (from each process to each one). After that, the preimages $x^{q+iS},\; 0\leq q<S,\; q\neq s$ of the points received by the $s$-th process from the neighbor ones are determined with the use of the evolvent $y^s(x)$. The points blocked within the $s$-th process will correspond to the preimages obtained. Then, each process performs the trials at the non-blocked points, the computations are performed in parallel using OpenMP. The results of the executed trials (the index of the point, the computed values of the problem functions, and the attribute of unblocking of this point) are transferred to all rest processes. All the points are added to the search information database, and the transition to the next iteration is performed.

@@

%GKLS generator for the functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.) has been proposed in \cite{Gaviano2003}. Four GKLS classes of differentiable test functions of the dimensions $N = 4$ and $5$, have been used. For each dimension, both \textit{Hard} and \textit{Simple} classes have been considered. The difficulty of a class was increased either by decreasing the radius of the attraction region of the global minimizer, or by decreasing the distance from the global minimizer $y^*$ to the domain boundaries. Application of the generator for studying some optimization algorithms has been described in \cite{Paulavicius2014,Sergeyev2015}.

In this study, we will use a GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce constrained problems. The scheme for generating constrained global optimization problems is proposed in \cite{Gergel2017}. Previous investigations have confirmed experimentally that the index method is not inferior to well-known analogues. A comparison of the method with the well known DIRECT method \cite{Jones2001} in solving unconstrained optimization problems has been performed in \cite{Barkalov2015}. In the present study, an experimental investigation is carried out of the speedup, which is obtained by the use of the index method in combination with the two-level parallelization scheme from \cite{Sidorov2015}.

The experiments have been carried out by solving a series of $100$ problems with two constraints and the objective functions from the \textit{Simple} and \textit{Hard} GKLS classes with the dimensionalities $N=4$, $N=5$. The number of the used cluster nodes $S$ and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at each node have been varied. The problem was considered to be solved, if the algorithm generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta$. The size of the vicinity was selected as $\delta=0.03\left\|b-a\right\|$, where $a$ and $b$ are the boundaries of the search domain. For the purpose of simulation of the computational complexity inherent to applied optimization problems, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (a series summation of $80$ thousand elements).

@@

\begin{table}
	\caption{Average time}
	\label{tab:1}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
		\multirow{2}{*}{\textit{Node(core)}} & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1(1)	&	220.5	&	334.8	&	&	1223.6	&	1386.6	\\
1(16)	&	31.3	&	49.1	&	&	211.8	&	547.2	\\
2(1)	&	158.4	&	260	&	&	1052.9	&	1458.1	\\
2(16)	&	22.1	&	35.9	&	&	227.5	&	603	\\
4(1)	&	127.7	&	286.4	&	&	951.3	&	1362.2	\\
4(16)	&	20.9	&	45	&	&	206	&	925.7	\\
8(1)	&	99.3	&	141.8	&	&	700.1	&	897.3	\\
8(16)	&	31	&	77.7	&	&	264.6	&	374	\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Average number of iterations}
	\label{tab:2}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
		\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1(1)	&	58320	&	84546	&	&	266943	&	287102	\\
1(16)	&	4297	&	6601	&	&	22655	&	56754	\\
2(1)	&	34791	&	52126	&	&	188465	&	241369	\\
2(16)	&	2029	&	3239	&	&	16689	&	40763	\\
4(1)	&	22223	&	47771	&	&	135734	&	180489	\\
4(16)	&	1281	&	2483	&	&	9241	&	35024	\\
8(1)	&	13844	&	18933	&	&	77748	&	94563	\\
8(16)	&	608	&	1473	&	&	5820	&	23033	\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

The results demonstrate the speedup achieved when using the shared memory at a node (performing several trials within a problem in parallel) as well as the distributed memory (solving in parallel several subproblems at different nodes). In this case, the highest time speedup was 10 (when using 64 cores on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster nodes). The difference in the speedups in time and in the number of iterations can be explained by the effect of the overheads of the data transmission between the processes. Note that when solving applied optimization problems, the computation of the problem function values even in one point is a computationally costly operation. In this case, the data transfer overheads will not affect significantly the total computational costs, and the difference between the time speedup and the iteration speedup will not be so great.
%Описание инфраструктуры

@@

The parallel algorithm has demonstrated a speedup with respect to the number of processors/cores employed. This was confirmed by the results of solving numerically of several hundred test problems using 128 cores of the Lobachevsky supercomputer. 
%The direction of further research is the generalization of the considered parallel algorithm for solving multicriteria problems.

@@

\bibitem{Balandin2011}
Balandin, D.V., Kogan, M.M.: Optimal linear-quadratic control: From matrix equations to linear matrix inequalities. Automation and Remote Control. 72(11), 2276--2284 (2011)

@@

\bibitem{Gergel2017}
Gergel, V.: An approach for generating test problems of constrained global optimization. Lecture Notes in Computer Science. 10556, 314--319 (2017)

\bibitem{Barkalov2015}
Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi coprocessor for solving global optimization problems. Lecture Notes in Computer Science. 9251, 307--318 (2015)

\bibitem{Sidorov2015}
Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of computationally intensive multiextremal optimization problems. Lecture Notes in Computer Science. 9251, 505--515 (2015)

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn

\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}

\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{multirow}
%
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
 \journalname{The Journal of Supercomputing}
%
\begin{document}

@@

The aim of this paper is to examine parallel methods for solving global optimization problems with non-convex constraints. The objective function and constraints are assumed to satisfy the Lipschitz condition with a priori unknown Lipschitz constants. The analytical form of the problem's functions may also be unknown, i.e. they can be set by an algorithm computing their values at various points within the search domain (the so-called ``black-box'' functions). It is also supposed that even a single computation of the function value can be time-consuming, since it requires performing numerical simulation in applied problems.

@@

\Russian
Данная ситуация является типичной для многих задач оптимального проектирования, когда при нарушении некоторых условий функционирования моделируемого объекта другие его характеристики оказываются неопределенными.

Among the examples of such problems are the optimal control problems described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2017}. It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e. every eigenvalue of $A$ has a strictly negative real part. Otherwise, the value of the criteria is indeterminate.

@@

\Russian
В статье \cite{Barkalov2017} приведено сравнение двух разных способа учета ограничений - классического метода штрафных функций и индексной схемы. Экспериментально подтрвеждено превосходство индексного метода над методом штрафных функций как по числу итераций, так и по числу вычсислений ограничений и целевой функции. Для решения задачи с заданной точностью индексному методу требовалось в 3-5 раз меньше итераций (в зависимости от размерности задачи и доли допусимой области). Одновременно с этим значение целевой функции вычислялось в среднем лишь на каждой пятой итерации индексного метода, в отличие от метода штрафных функций, в котором значения целевой функции вычисляются на каждой итерации.

@@

Another scheme of building of the multiple mapping was proposed  in \cite{6_Gergel2009}. The building of a set of Peano curves not by the shift along the main diagonal of the hypercube but by rotation of the evolvents around the coordinate origin is a distinctive feature of the proposed scheme. In the initial non-rotated mapping for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other. In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer. The maximum number of various rotations of the evolvents mapping the $N$-dimensional hypercube onto a one-dimensional interval is $2^N$. The employment of all possible rotations might appear to be redundant. In this case, we can select only a part of all rotations. As a possible approach, we propose to generate the new evolvents by the rotation of the initial evolvent on the angle of $\pm\pi/2$ in each of the coordinate planes. Taking into account the initial evolvent the total number of such evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals equals to $N(N-1)+1$. 
\Russian
Именно такой подход к построению multiple mapping был использован в данной статье при проведении вычислительных экспериментов

@@

@@

In this study, we will use a GKLS generator of the functions of arbitrary dimensionality proposed in \cite{Gaviano2003} to produce constrained problems. The scheme for generating constrained global optimization problems is proposed in \cite{Gergel2017}. 
\Russian
Тестовые задачи с ограничениями, полученные с помощью данной схемы, обладают важным свойством, присущим прикладным задачам оптимизации: условный глобальный минимум находится на границе допустимой области, в то время как глобальный минимум целевой функции без ограничений находится вне допустимой области. При этом можно варьировать объем допустимой области, а также число ограничений, активных в точке минимума.

%Previous investigations have confirmed experimentally that the index method is not inferior to well-known analogues. A comparison of the method with the well known DIRECT method \cite{Jones2001} in solving unconstrained optimization problems has been performed in \cite{Barkalov2015}. 
In the present study, an experimental investigation is carried out of the speedup, which is obtained by the use of the index method in combination with the two-level parallelization scheme from \cite{Sidorov2015}.

The experiments have been carried out by solving a series of $100$ problems with two constraints and one objective function from the \textit{Simple} and \textit{Hard} GKLS classes with the dimensionalities $N=4$, $N=5$. The number of the used cluster nodes $S$ and, correspondingly, the number of evolvents as well as the number of cores $p$ employed at each node have been varied. The problem was considered to be solved, if the algorithm generated a trial point $y^k$ in the $\delta$-vicinity of the global minimum, i.e., $\left\|y^k-y^*\right\|\leq \delta$. The size of the vicinity was selected as $\delta=0.03\left\|b-a\right\|$, where $a$ and $b$ are the boundaries of the search domain. For the purpose of simulation of the computational complexity inherent to applied optimization problems, calculation of the problem functions in all performed experiments was made more complex by additional calculations without changing the type of function and arrangement of its minima (a series summation of $100$ thousand elements).

@@

\begin{table}
	\caption{Average time}
	\label{tab:1}
	\center
	\begin{tabular}{cccccc}
		\hline\noalign{\smallskip}
		\multirow{2}{*}{\textit{Node(core)}} & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1(1)	&	220.5	&	334.8	&	&	1223.6	&	1386.6	\\
1(16)	&	31.3	&	49.1	&	&	211.8	&	547.2	\\
2(1)	&	158.4	&	260	&	&	1052.9	&	1458.1	\\
2(16)	&	22.1	&	35.9	&	&	227.5	&	603	\\
4(1)	&	127.7	&	286.4	&	&	951.3	&	1362.2	\\
4(16)	&	20.9	&	45	&	&	206	&	925.7	\\
8(1)	&	99.3	&	141.8	&	&	700.1	&	897.3	\\
8(16)	&	31	&	77.7	&	&	264.6	&	374	\\
12(1)	&	99.3	&	141.8	&	&	700.1	&	897.3	\\
12(16)	&	31	&	77.7	&	&	264.6	&	374	\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

%\begin{table}
	%\caption{Average number of iterations}
	%\label{tab:2}
	%\center
	%\begin{tabular}{cccccc}
		%\hline\noalign{\smallskip}
		%\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$} \\
		%\noalign{\smallskip} \cline{2-3} \cline{5-6} \noalign{\smallskip}
		 %& \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		%\noalign{\smallskip} \hline \noalign{\smallskip}
%1(1)	&	58320	&	84546	&	&	266943	&	287102	\\
%1(16)	&	4297	&	6601	&	&	22655	&	56754	\\
%2(1)	&	34791	&	52126	&	&	188465	&	241369	\\
%2(16)	&	2029	&	3239	&	&	16689	&	40763	\\
%4(1)	&	22223	&	47771	&	&	135734	&	180489	\\
%4(16)	&	1281	&	2483	&	&	9241	&	35024	\\
%8(1)	&	13844	&	18933	&	&	77748	&	94563	\\
%8(16)	&	608	&	1473	&	&	5820	&	23033	\\
%12(1)	&	13844	&	18933	&	&	77748	&	94563	\\
%12(16)	&	608	&	1473	&	&	5820	&	23033	\\
		%\noalign{\smallskip}\hline
	%\end{tabular}
%\end{table}

\begin{table}
	\caption{Average number of iterations}
	\label{tab:2}
	\center
	\begin{tabular}{ccccccccc}
		\hline\noalign{\smallskip}
		\multirow{2}{*}{\textit{Node(core)}}	 & \multicolumn{2}{c}{ $N=4$ } & & \multicolumn{2}{c}{$N=5$}& & \multicolumn{2}{c}{$N=6$} \\
		\noalign{\smallskip} \cline{2-3} \cline{5-6} \cline{8-9} \noalign{\smallskip}
		 & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard} & & \textit{Simple} & \textit{Hard}  \\
		\noalign{\smallskip} \hline \noalign{\smallskip}
1(1)	&	58320	&	84546	&	&	266943	&	287102 & & 1 & 2	\\
1(16)	&	4297	&	6601	&	&	22655	&	56754	& &  & \\
2(1)	&	34791	&	52126	&	&	188465	&	241369	& &  &\\
2(16)	&	2029	&	3239	&	&	16689	&	40763	& &  &\\
4(1)	&	22223	&	47771	&	&	135734	&	180489 & &  &	\\
4(16)	&	1281	&	2483	&	&	9241	&	35024	& &  &\\
8(1)	&	13844	&	18933	&	&	77748	&	94563	& &  &\\
8(16)	&	608	&	1473	&	&	5820	&	23033	& &  &\\
12(1)	&	13844	&	18933	&	&	77748	&	94563	& &  &\\
12(16)	&	608	&	1473	&	&	5820	&	23033	& &  &\\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

The results demonstrate the speedup achieved when using the shared memory at a node (performing several trials within a problem in parallel) as well as the distributed memory (solving in parallel several subproblems at different nodes). In this case, the highest time speedup was 10 (when using 64 cores on 4 cluster nodes), the highest iteration speedup was 95 (when using 128 cores on 8 cluster nodes). The difference in the speedups in time and in the number of iterations can be explained by the effect of the overheads of the data transmission between the processes. Note that when solving applied optimization problems, the computation of the problem function values even in one point is a computationally costly operation. In this case, the data transfer overheads will not affect significantly the total computational costs, and the difference between the time speedup and the iteration speedup will not be so great.

@@

The parallel algorithm has demonstrated a speedup with respect to the number of processors/cores employed. This was confirmed by the results of solving numerically of several hundred test problems using 128 cores of the Lobachevsky supercomputer. 
\Russian
Что-то про применение алгоритма. 
%The direction of further research is the generalization of the considered parallel algorithm for solving multicriteria problems.

@@

%\bibitem{Balandin2011} Balandin, D.V., Kogan, M.M.: Optimal linear-quadratic control: From matrix equations to linear matrix inequalities. Automation and Remote Control. 72(11), 2276--2284 (2011)

@@

\bibitem{Gergel2017}
Gergel, V.: An approach for generating test problems of constrained global optimization. LNCS. 10556, 314--319 (2017)

%\bibitem{Barkalov2015} Barkalov, K., Gergel, V., Lebedev, I.: Use of Xeon Phi coprocessor for solving global optimization problems. LNCS. 9251, 307--318 (2015)

\bibitem{Sidorov2015}
Gergel, V., Sidorov, S.: A two-level parallel global search algorithm for solution of computationally intensive multiextremal optimization problems. LNCS. 9251, 505--515 (2015)

\bibitem{Barkalov2017}
Barkalov, K., Lebeled, I.: Comparing Two Approaches for Solving Constrained Global Optimization Problems. LNCS. 10556, 301--306 (2017)

\bibitem{6_Gergel2009}
Strongin, R.G., Gergel, V.P., Barkalov, K.A.: Parallel methods for global optimization problem solving. Journal of instrument engineering. 52, 25--33 (2009) (In Russian)
